A spatial analysis of the environmental correlates of COVID-19 incidence in Spain

### 2020-04-18 Environmental-Correlates-of-COVID19-Spain/Environmental-Correlates-of-COVID19-Spain-bkp.Rmd
Created backup file

```{r formulas, include=FALSE, cache = TRUE}
# Define formulas with three different lagged variables:
formula_lag8 <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag8) +
  log(Mean_Temp_lag8)

formula_lag11 <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag11) +
  log(Mean_Temp_lag11)

formula_lag11w <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag11w) +
  log(Mean_Temp_lag11w)
```

```{r model-restrictions, include=FALSE, cache = TRUE}
T <- max(covid19_spain$Date) - min(covid19_spain$Date) + 1 # Recall that T is the number of days, i.e., time periods, i.e., equations; add 1 to include the starting day
k <- 7 # Number of independent variables, including the constant
coef_rest <- 2 # Number of restrictions

@@

for (i in 1:(T-1)){
  R2[i, 2] <- 1
  R2[i, (2 + i * k)] <- -1
  R2[(i + T - 1), 3] <- 1
  R2[(i + T - 1), (3 + i * k)] <- -1
  # Use if more restrictions are needed
  #R2[(i + T - 1) * 2, 4] <- 1
  #R2[(i + T - 1) * 2, (4 + i * k)] <- -1
}
b2 <- matrix(0, ncol = 21*coef_rest)
```

@@

Summary of best model.
```{r summary-best-model, echo=FALSE, cache = TRUE}
data.frame(Coefficient = sur.slm_lag11$coefficients, 
           pval = dnorm(abs(sur.slm_lag11$coefficients/sur.slm_lag11$rest.se))) %>%
  rownames_to_column(var = "Variable") %>%
  mutate(Variable = c("Intercept", 
                      "log(GDPpc)", 
                      "log(Older)",
                      "log(Density)",
                      "Transit",
                      "log(Humidity)",
                      "log(Temperature)",
                      rep(c("Intercept",
                            "log(Density)", 
                            "Transit", 
                            "log(Humidity)", 
                            "log(Temperature)"), 29)),
         Variable = factor(Variable, 
                           levels = c("Intercept", 
                                      "log(GDPpc)", 
                                      "log(Older)", 
                                      "log(Density)",
                                      "Transit", 
                                      "log(Humidity)",
                                      "log(Temperature)"))) %>%
  group_by(Variable) %>%
  summarize(Min = min(Coefficient), 
            Mean = mean(Coefficient), 
            Max = max(Coefficient),
            pns = sum(pval > 0.10), # Number of coefficients with p > 0.1
            p10 = sum(pval <= 0.10) - sum(pval <= 0.05), # Number of coefficients with  0.1 <= pval < 0.05
            p05 = sum(pval <= 0.05)) %>% # Number of coefficients with pval <=5 
  mutate(Note = c("Non-constrained", 
                  "Constrained", 
                  "Constrained", 
                  "Non-constrained", 
                  "Non-constrained",
                  "Non-constrained",
                  "Non-constrained")) %>%
  kable("latex",
        #"html",
        booktabs = TRUE,
        digits = 3,
        col.names = c("Variable", "Min", "Mean", "Max", "p > 0.10", "0.10 <= p < 0.05", "p <= 0.05", "Note"),
        caption = "\\label{tab:summary-best-model}Summary of estimation results best model (lagged 11-day moving average)") %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  #kable_styling(bootstrap_options = c("striped", "condensed")) %>%
  add_header_above(c(" " = 1, "Estimates" = 3, "Significance" = 3, " " = 1)) %>%
  footnote(general = c("Significance: This is the number of coefficients with p-values as indicated",
                       "Non-constrained: coefficient was allowed to vary across equations",
                       "Constrained: coefficient as constant across equations"))
```

@@

Figure \ref{fig:beta-density-time} shows the temporal evolution of the coefficient for $\log(Temperature)$.

@@

>>>>

```{r formulas, include=FALSE, cache = TRUE}
# Define formulas with three different lagged variables:
formula_lag8 <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag8) +
  log(Mean_Temp_lag8) +
  log(Sunshine_Hours_lag8 + 0.1)

formula_lag11 <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag11) +
  log(Mean_Temp_lag11) +
  log(Sunshine_Hours_lag11 + 0.1)

formula_lag11w <- log(Incidence) ~ 
  log(GDPpc) +
  log(Older) +
  log(Density) +
  Transit +
  log(Humidity_lag11w) +
  log(Mean_Temp_lag11w) +
  log(Sunshine_Hours_lag11w + 0.1)
```

```{r model-restrictions, include=FALSE, cache = TRUE}
T <- max(covid19_spain$Date) - min(covid19_spain$Date) + 1 # Recall that T is the number of days, i.e., time periods, i.e., equations; add 1 to include the starting day
k <- 8 # Number of independent variables, including the constant
coef_rest <- 2 # Number of restrictions

@@

for (i in 1:(T-1)){
  R2[i, 2] <- 1
  R2[i, (2 + i * k)] <- -1
  R2[(i + T - 1), 3] <- 1
  R2[(i + T - 1), (3 + i * k)] <- -1
  # Use if more restrictions are needed
  #R2[(i + T - 1) * 2, 4] <- 1
  #R2[(i + T - 1) * 2, (4 + i * k)] <- -1
}
b2 <- matrix(0, ncol = (T - 1) * coef_rest)
```

@@

Summary of best model.
```{r summary-best-model, echo=FALSE, cache = TRUE}
data.frame(Coefficient = sur.slm_lag11$coefficients, 
           pval = dnorm(abs(sur.slm_lag11$coefficients/sur.slm_lag11$rest.se))) %>%
  rownames_to_column(var = "Variable") %>%
  mutate(Variable = c("Intercept", 
                      "log(GDPpc)", 
                      "log(Older)",
                      "log(Density)",
                      "Transit",
                      "log(Humidity)",
                      "log(Temperature)",
                      "log(Sunshine)",
                      rep(c("Intercept",
                            "log(Density)", 
                            "Transit", 
                            "log(Humidity)", 
                            "log(Temperature)",
                            "log(Sunshine)"),
                          29)),
         Variable = factor(Variable, 
                           levels = c("Intercept", 
                                      "log(GDPpc)", 
                                      "log(Older)", 
                                      "log(Density)",
                                      "Transit", 
                                      "log(Humidity)",
                                      "log(Temperature)",
                                      "log(Sunshine)"))) %>%
  group_by(Variable) %>%
  summarize(Min = min(Coefficient), 
            Mean = mean(Coefficient), 
            Max = max(Coefficient),
            pns = sum(pval > 0.10), # Number of coefficients with p > 0.1
            p10 = sum(pval <= 0.10) - sum(pval <= 0.05), # Number of coefficients with  0.1 <= pval < 0.05
            p05 = sum(pval <= 0.05)) %>% # Number of coefficients with pval <=5 
  mutate(Note = c("Non-constrained", 
                  "Constrained", 
                  "Constrained", 
                  "Non-constrained", 
                  "Non-constrained",
                  "Non-constrained",
                  "Non-constrained",
                  "Non-constrained")) %>%
  kable("latex",
        #"html",
        booktabs = TRUE,
        digits = 3,
        col.names = c("Variable", "Min", "Mean", "Max", "p > 0.10", "0.10 <= p < 0.05", "p <= 0.05", "Note"),
        caption = "\\label{tab:summary-best-model}Summary of estimation results best model (lagged 11-day moving average)") %>%
  kable_styling(latex_options = c("striped", "scale_down")) %>%
  #kable_styling(bootstrap_options = c("striped", "condensed")) %>%
  add_header_above(c(" " = 1, "Estimates" = 3, "Significance" = 3, " " = 1)) %>%
  footnote(general = c("Significance: This is the number of coefficients with p-values as indicated",
                       "Non-constrained: coefficient was allowed to vary across equations",
                       "Constrained: coefficient as constant across equations"))
```

@@

Figure \ref{fig:beta-temperature-time} shows the temporal evolution of the coefficient for $\log(Temperature)$.

@@

Figure \ref{fig:beta-sunshine-time} shows the temporal evolution of the coefficient for $\log(Sunshine + 0.1)$.

```{r beta-sunshine-time, echo=FALSE, fig.cap="\\label{fig:beta-sunshine-time}Temporal evolution of coefficient for log(Sunshine + 0.1)", cache = TRUE}
n = 6
data.frame(Date = seq(ymd("2020-03-13"), 
                      ymd("2020-04-11"), 
                      by = "days"),
           Beta = matrix(sur.slm_lag11$coefficients[-c(2,3)],ncol = T)[n,],
           tvalue = matrix(sur.slm_lag11$coefficients[-c(2,3)]/sur.slm_lag11$rest.se[-c(2,3)], ncol = T)[n,]) %>%
  mutate(sig = case_when(abs(tvalue) >= qnorm(0.95, mean = 0, sd = 1) ~ "p <= 0.05",
                         abs(tvalue) >= qnorm(0.90, mean = 0, sd = 1) ~ "p <= 0.10",
                         TRUE ~ "p > 0.10")) %>%
  ggplot(aes(x = Date, y = Beta, color = sig)) +
  geom_point(size = 2) +
  scale_color_manual(values = c("p <= 0.10" = "orange", "p <= 0.05" = "red", "p > 0.10" = "black")) +  geom_hline(yintercept = 0) +
  geom_vline(xintercept = as_date("2020-03-16"), color = "red") +
  theme_tufte()
```

### 2020-04-18 Environmental-Correlates-of-COVID19-Spain/Environmental-Correlates-of-COVID19-Spain.tex
Created backup file

https://github.com/paezha/covid19-environmental-correlates

@@

\#\# Covid-19 in Spain

@@

\hypertarget{sources-of-data}{%
\subsection{Sources of Data}\label{sources-of-data}}

@@

This specification assumes that incidence in a province (\(y_st\)) at
time \(t\) is partially determined by the weighted average (\(Wy_{st}\))
of incidence in neighbouring provinces. Coefficients for the spatially
lagged variable are estimated for each time period \(\rho_t\) and
identifies the intensity and the sign of the impacts of neighbourhood.
It is possible test the null hypotheis of identtical levels of spatial
dependence (\(\rho_i=\rho_j, \forall i,j\)). The correspond Wald test is
available in the R package \texttt{spsur}.

@@

\caption{\label{tab:summary-best-model}\label{tab:summary-best-model}Summary of estimation results best model (lagged 11-day moving average)}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Estimates} & \multicolumn{3}{c}{Significance} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
Variable & Min & Mean & Max & p > 0.10 & 0.10 <= p < 0.05 & p <= 0.05 & Note\\
\midrule
\rowcolor{gray!6}  Intercept & 7.175 & 10.008 & 13.845 & 0 & 0 & 30 & Non-constrained\\
log(GDPpc) & 0.513 & 0.513 & 0.513 & 0 & 1 & 0 & Constrained\\
\rowcolor{gray!6}  log(Older) & -0.783 & -0.783 & -0.783 & 0 & 0 & 1 & Constrained\\
log(Density) & -0.217 & -0.092 & 0.188 & 19 & 1 & 10 & Non-constrained\\
\rowcolor{gray!6}  Transit & 0.300 & 0.518 & 0.595 & 11 & 10 & 9 & Non-constrained\\
\addlinespace
log(Humidity) & -1.176 & -0.372 & 0.342 & 17 & 4 & 9 & Non-constrained\\
\rowcolor{gray!6}  log(Temperature) & -2.048 & -1.431 & -0.989 & 0 & 0 & 30 & Non-constrained\\
\bottomrule
\multicolumn{8}{l}{\textit{Note: }}\\
\multicolumn{8}{l}{Significance: This is the number of coefficients with p-values as indicated}\\
\multicolumn{8}{l}{Non-constrained: coefficient was allowed to vary across equations}\\
\multicolumn{8}{l}{Constrained: coefficient as constant across equations}\\
\end{tabular}}
\end{table}

@@

Figure \ref{fig:delta-time} shows the temporal evolution of the spatial
autocorrelation coefficient (\(\rho\)).

@@

Figure \ref{fig:beta-density-time} shows the temporal evolution of the
coefficient for \(\log(Temperature)\).

@@

>>>>

\begin{quote}
https://github.com/paezha/covid19-environmental-correlates
\end{quote}

@@

\hypertarget{covid-19-in-spain}{%
\subsection{Covid-19 in Spain}\label{covid-19-in-spain}}

@@

\hypertarget{data}{%
\subsection{Data}\label{data}}

@@

This specification assumes that incidence in a province (\(y_st\)) at
time \(t\) is partially determined by the weighted average (\(Wy_{st}\))
of incidence in neighbouring provinces. Coefficients for the spatially
lagged variable are estimated for each time period \(\lambda_t\) and
identifies the intensity and the sign of the impacts of neighbourhood.
It is possible test the null hypotheis of identtical levels of spatial
dependence (\(\lambda_i=\lambda_j, \forall i,j\)). The correspond Wald
test is available in the \texttt{R} package \texttt{spsur}.

@@

\caption{\label{tab:summary-best-model}\label{tab:summary-best-model}Summary of estimation results best model (lagged 11-day moving average)}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{lrrrrrrl}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Estimates} & \multicolumn{3}{c}{Significance} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7}
Variable & Min & Mean & Max & p > 0.10 & 0.10 <= p < 0.05 & p <= 0.05 & Note\\
\midrule
\rowcolor{gray!6}  Intercept & 7.370 & 10.037 & 12.968 & 0 & 0 & 30 & Non-constrained\\
log(GDPpc) & 0.620 & 0.620 & 0.620 & 0 & 0 & 1 & Constrained\\
\rowcolor{gray!6}  log(Older) & -0.737 & -0.737 & -0.737 & 0 & 0 & 1 & Constrained\\
log(Density) & -0.220 & -0.097 & 0.153 & 19 & 1 & 10 & Non-constrained\\
\rowcolor{gray!6}  Transit & 0.314 & 0.512 & 0.583 & 10 & 9 & 11 & Non-constrained\\
\addlinespace
log(Humidity) & -1.434 & -0.534 & -0.031 & 10 & 1 & 19 & Non-constrained\\
\rowcolor{gray!6}  log(Temperature) & -2.014 & -1.406 & -0.929 & 0 & 0 & 30 & Non-constrained\\
log(Sunshine) & -0.258 & 0.097 & 0.206 & 7 & 2 & 21 & Non-constrained\\
\bottomrule
\multicolumn{8}{l}{\textit{Note: }}\\
\multicolumn{8}{l}{Significance: This is the number of coefficients with p-values as indicated}\\
\multicolumn{8}{l}{Non-constrained: coefficient was allowed to vary across equations}\\
\multicolumn{8}{l}{Constrained: coefficient as constant across equations}\\
\end{tabular}}
\end{table}

@@

Figure \ref{fig:delta-time} shows the temporal evolution of the spatial
autocorrelation coefficient (\(\lambda\)).

@@

Figure \ref{fig:beta-temperature-time} shows the temporal evolution of
the coefficient for \(\log(Temperature)\).

@@

Figure \ref{fig:beta-sunshine-time} shows the temporal evolution of the
coefficient for \(\log(Sunshine + 0.1)\).

\begin{figure}
\centering
\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/beta-sunshine-time-1.pdf}
\caption{\label{fig:beta-sunshine-time}Temporal evolution of coefficient
for log(Sunshine + 0.1)}
\end{figure}
